Spatial phylogenetics of Japanese ferns: Patterns, processes, and implications for conservation

### 2021-09-21 ms/SI.Rmd
Update MS

---
title: "Appendix S1: Supporting Tables and Figures"
mainfont: Roboto-Regular.ttf
fontsize: 11pt
mainfontoptions: 
  - BoldFont=Roboto-Bold.ttf
  - ItalicFont=Roboto-Italic.ttf
  - BoldItalicFont=Roboto-BoldItalic.ttf
output: 
  bookdown::pdf_document2:
    citation_package: default
    toc: no
    number_sections: no
    fig_caption: yes
    keep_tex: no
    latex_engine: xelatex
    pandoc_args: [ "--csl", "apa-6th-edition.csl"]
header-includes: 
  - \usepackage{float} # keep figures from floating away
  - \usepackage{booktabs} # need for kable tables
  - \usepackage{colortbl} # need for kable tables
  - \makeatletter\renewcommand*{\fps@figure}{H}\makeatother # keep figures from floating away
  - \usepackage{fancyhdr} # enable header
  - \pagestyle{fancy}  # enable header
  - \fancypagestyle{plain}{\pagestyle{fancy}} # show header on first page
  - \renewcommand{\headrulewidth}{0pt} # delete line beneath header
  - \fancyhead[RE,RO]{Nitta \textit{et al}. Spatial phy. of Japanese ferns Appendix S1} # add header on right side
  - \fancyhead[LO,LE]{} # suppress automatic section header on left side
  - \cfoot{\thepage} # footer
editor_options: 
  chunk_output_type: console
params:
  doc_type: pdf
# nocite: | 
---

@@

**`r s_figure("sampling-curve")`**. Species collection curve for the ferns of Japan fit with the `r pack("iNEXT")` package [@Hsieh2016]. Grey shading indicates 95% confidence interval (not visible where it does not extend beyond the fitted line). Point indicates observed taxonomic richness (`r observed_data$qD` taxa, 95% confidence interval `r rich_interval` taxa). Taxa include species, subspecies, and varieties, excluding hybrids. Sampling completeness is estimated to be `r est_completeness` (95% confidence interval `r completeness_interval`).

>>>>

---
title: "Appendix S1: Supporting Tables and Figures"
mainfont: Roboto-Regular.ttf
fontsize: 11pt
mainfontoptions: 
  - BoldFont=Roboto-Bold.ttf
  - ItalicFont=Roboto-Italic.ttf
  - BoldItalicFont=Roboto-BoldItalic.ttf
output: 
  bookdown::pdf_document2:
    citation_package: default
    toc: no
    number_sections: no
    fig_caption: yes
    keep_tex: no
    latex_engine: xelatex
    pandoc_args: [ "--csl", "apa-6th-edition.csl"]
header-includes: 
  - \usepackage{float} # keep figures from floating away
  - \usepackage{booktabs} # need for kable tables
  - \usepackage{colortbl} # need for kable tables
  - \makeatletter\renewcommand*{\fps@figure}{H}\makeatother # keep figures from floating away
  - \usepackage{fancyhdr} # enable header
  - \pagestyle{fancy}  # enable header
  - \fancypagestyle{plain}{\pagestyle{fancy}} # show header on first page
  - \renewcommand{\headrulewidth}{0pt} # delete line beneath header
  - \fancyhead[RE,RO]{Nitta et al.---\textit{American Journal of Botany} 2021 -- Appendix S1} # add header on right side
  - \fancyhead[LO,LE]{} # suppress automatic section header on left side
  - \cfoot{\thepage} # footer
editor_options: 
  chunk_output_type: console
params:
  doc_type: pdf
# nocite: | 
---

@@

**`r s_figure("sampling-curve")`**. Species collection curve for the ferns of Japan. Curve fit with the `r pack("iNEXT")` package [@Hsieh2016]. Grey shading indicates 95% confidence interval (not visible where it does not extend beyond the fitted line). Point indicates observed taxonomic richness (`r observed_data$qD` taxa, 95% confidence interval `r rich_interval` taxa). Taxa include species, subspecies, and varieties, excluding hybrids. Sampling completeness is estimated to be `r est_completeness` (95% confidence interval `r completeness_interval`).

### 2021-09-21 ms/data_exploration.Rmd
Update MS

---
title: "Appendix S2: Data Exploration"
mainfont: Roboto-Regular.ttf
fontsize: 11pt
mainfontoptions: 
  - BoldFont=Roboto-Bold.ttf
  - ItalicFont=Roboto-Italic.ttf
  - BoldItalicFont=Roboto-BoldItalic.ttf
output: 
  bookdown::pdf_document2:
    citation_package: default
    toc: no
    number_sections: no
    fig_caption: yes
    keep_tex: no
    latex_engine: xelatex
    pandoc_args: [ "--csl", "apa-6th-edition.csl"]
editor_options: 
  chunk_output_type: console
header-includes: 
  - \usepackage{float} # keep figures from floating away
  - \makeatletter\renewcommand*{\fps@figure}{H}\makeatother # keep figures from floating away
  - \usepackage{fancyhdr} # enable header
  - \pagestyle{fancy}  # enable header
  - \fancypagestyle{plain}{\pagestyle{fancy}} # show header on first page
  - \renewcommand{\headrulewidth}{0pt} # delete line beneath header
  - \fancyhead[RE,RO]{Nitta \textit{et al}. Spatial phy. of Japanese ferns Appendix S2} # add header on right side
  - \fancyhead[LO,LE]{} # suppress automatic section header on left side
  - \cfoot{\thepage} # footer
---

@@

Here we perform initial data exploration to determine the appropriate models to fit to the data.

@@

```{r load-data, message = FALSE}
# Load the data from the targets workflow
# tar_load(biodiv_ferns_cent_dryad)

# Alternatively, unzip `results.zip` in the Dryad data
# (FIXME: add DOI when available)
# and read in the data from there
biodiv_ferns_cent_dryad <- read_csv(here::here("results/dryad_files/japan_ferns_biodiv.csv"))
```

>>>>

---
title: "Appendix S2: Exploratory Data Analysis"
mainfont: Roboto-Regular.ttf
fontsize: 11pt
mainfontoptions: 
  - BoldFont=Roboto-Bold.ttf
  - ItalicFont=Roboto-Italic.ttf
  - BoldItalicFont=Roboto-BoldItalic.ttf
output: 
  bookdown::pdf_document2:
    citation_package: default
    toc: no
    number_sections: no
    fig_caption: yes
    keep_tex: no
    latex_engine: xelatex
    pandoc_args: [ "--csl", "apa-6th-edition.csl"]
editor_options: 
  chunk_output_type: console
header-includes: 
  - \usepackage{float} # keep figures from floating away
  - \makeatletter\renewcommand*{\fps@figure}{H}\makeatother # keep figures from floating away
  - \usepackage{fancyhdr} # enable header
  - \pagestyle{fancy}  # enable header
  - \fancypagestyle{plain}{\pagestyle{fancy}} # show header on first page
  - \renewcommand{\headrulewidth}{0pt} # delete line beneath header
  - \fancyhead[RE,RO]{Nitta et al.---\textit{American Journal of Botany} 2021 -- Appendix S2} # add header on right side
  - \fancyhead[LO,LE]{} # suppress automatic section header on left side
  - \cfoot{\thepage} # footer
---

@@

Here we perform exploratory data analysis to determine the appropriate models to fit to the data.

@@

```{r load-data, message = FALSE}
# Load the data from the targets workflow
tar_load(biodiv_ferns_cent_dryad)

# Alternatively, unzip `results.zip` in the Dryad data
# (FIXME: add DOI when available)
# and read in the data from there
# biodiv_ferns_cent_dryad <- read_csv(here::here("japan_ferns_biodiv.csv"))
```

### 2021-09-21 ms/manuscript.Rmd
Update MS

---
title: "Spatial phylogenetics of Japanese ferns: Patterns, processes, and implications for conservation"
mainfont: Roboto-Regular.ttf
fontsize: 11pt
mainfontoptions: 
  - BoldFont=Roboto-Bold.ttf
  - ItalicFont=Roboto-Italic.ttf
  - BoldItalicFont=Roboto-BoldItalic.ttf
output: 
  bookdown::word_document2:
    pandoc_args: [ "--csl", "apa-6th-edition.csl"]
    toc: no
    number_sections: no
    reference_docx: ajb_template.docx
  bookdown::pdf_document2:
    citation_package: default
    toc: no
    number_sections: no
    fig_caption: no
    latex_engine: xelatex
    pandoc_args: [ "--csl", "apa-6th-edition.csl"]
params:
  doc_type: doc
editor_options: 
  chunk_output_type: console
---

@@

## Abstract

@@

## MATERIALS AND METHODS

@@

### Study Site

Our data include all ferns known to natively occur within the boundaries of Japan (20°N to 45°N, 122°E to 153°E). 
Japan consists of four main islands and thousands of smaller ones (`r figure("japan-map")`). 
Most of the main islands have been in contact with the continent at various points in the past during periods of lower sea level, but the Ogasawara archipelago consists of oceanic islands that have never been connected to the mainland. 
The climate is subtropical in the south and temperate in other parts of the country. 
Map data for Japan were downloaded from the Geospatial Information Authority of Japan under the Creative Commons Attribution License v4.0 (https://www.gsi.go.jp/kankyochiri/gm_japan_e.html).

### Occurrence data

@@

### Morphological trait data

@@

### Reproductive mode data

@@

### Environmental data

@@

### Phylogenetic analysis

@@

### Phylogenetic signal

@@

### Analysis of biodiversity

@@

### Spatial models

@@

### Analysis of bioregions

@@

### Assessment of conservation status and threats

@@

## RESULTS

### Datasets

@@

### Observed diversity

@@

### Randomization tests for phylogenetic and functional diversity

@@

### Relationships between biodiversity and climate or reproductive mode

@@

### Categorical analysis of neo- and paleo-endemism

@@

### Bioregions

@@

### Conservation status and threats

@@

## DISCUSSION

### Distribution and drivers of biodiversity

@@

### CONCLUSIONS

@@

Our study integrates diverse, thoroughly collected datasets to characterize the spatial biodiversity of Japanese ferns and gain insight into the processes generating these patterns. We end by summarizing some caveats that must be kept in mind when interpreting our results, while also suggesting avenues for future research.

@@

## Acknowledgements

Members of the Iwasaki Lab at the University of Tokyo provided comments that improved the manuscript.
Shawn Laffan provided advice about CANAPE.
This study was supported by JSPS KAKENHI Grant Number 16H06279.

## Author contributions

@@

## Data availability

@@

## Literature cited
<!-- Insert <div id="refs"></div> to place the refs here instead of at the end -->
<div id="refs"></div>

@@

**`r figure("raw-diversity")`**. (a) Raw taxonomic diversity (b), phylogenetic diversity, and (c) functional diversity of the ferns of Japan. For (b) and (c), raw branchlengths were transformed to relative values by dividing each branchlength by the sum of all branchlengths.

>>>>

---
title: "Spatial phylogenetics of Japanese ferns: Patterns, processes, and implications for conservation"
mainfont: Roboto-Regular.ttf
fontsize: 11pt
mainfontoptions: 
  - BoldFont=Roboto-Bold.ttf
  - ItalicFont=Roboto-Italic.ttf
  - BoldItalicFont=Roboto-BoldItalic.ttf
output: 
  bookdown::word_document2:
    pandoc_args: [ "--csl", "american-journal-of-botany.csl"]
    toc: no
    number_sections: no
    reference_docx: ajb_template.docx
  bookdown::pdf_document2:
    citation_package: default
    toc: no
    number_sections: no
    fig_caption: no
    latex_engine: xelatex
    pandoc_args: [ "--csl", "american-journal-of-botany.csl"]
params:
  doc_type: doc
editor_options: 
  chunk_output_type: console
---

@@

**Abstract**

@@

# Materials and methods

@@

## Study Site

Our data include all ferns known to natively occur within the boundaries of Japan . 
Japan (20°N to 45°N, 122°E to 153°E) consists of four main islands and thousands of smaller ones (`r figure("japan-map")`). 
Most of the main islands have been in contact with the continent at various points in the past during periods of lower sea level, but the Ogasawara archipelago consists of oceanic islands that have never been connected to the mainland. 
The climate is subtropical in the south and temperate in other parts of the country. 
Map data for Japan were downloaded from the Geospatial Information Authority of Japan under the Creative Commons Attribution License v4.0 (https://www.gsi.go.jp/kankyochiri/gm_japan_e.html).

## Occurrence data

@@

## Morphological trait data

@@

## Reproductive mode data

@@

## Environmental data

@@

## Phylogenetic analysis

@@

## Phylogenetic signal

@@

## Analysis of biodiversity

@@

## Spatial models

@@

## Analysis of bioregions

@@

## Assessment of conservation status and threats

@@

# Results

## Datasets

@@

## Observed diversity

@@

## Randomization tests for phylogenetic and functional diversity

@@

## Relationships between biodiversity and climate or reproductive mode

@@

## Categorical analysis of neo- and paleo-endemism

@@

## Bioregions

@@

## Conservation status and threats

@@

# Discussion

## Distribution and drivers of biodiversity

@@

# Conclusions

@@

Our study integrates diverse, thoroughly collected datasets to characterize the spatial biodiversity of Japanese ferns and gain insight into the processes generating these patterns. We conclude by summarizing some caveats that must be kept in mind when interpreting our results, while also suggesting avenues for future research.

@@

# Acknowledgements

The authors thank Members of the Iwasaki Lab at the University of Tokyo for providing comments that improved the manuscript.
Shawn Laffan provided advice about CANAPE.
This study was supported by JSPS KAKENHI Grant Number 16H06279.

# Author contributions

@@

# Data availability statement

@@

# Online supporting information

Additional supporting information may be found online in the Supporting Information section at the end of the article.

**Appendix S1**. Supplementary tables and figures, including Tables `r s_table_num("moran")`--`r s_table_num("aic")` and Figs. `r s_figure_num("sampling-curve")`--`r s_figure_num("deer-map")`.

**`r s_table("moran")`**. Spatial autocorrelation in biodiversity metrics (richness, SES of PD, SES of RPD, SES of FD, SES of RFD) and independent variables (environmental variables and % apomictic taxa) as measured with Moran's *I*.  
**`r s_table("corr")`**. Results of modified *t*-test for correlation between variables while taking into account spatial position.  
**`r s_table("phy-sig")`**. Phylogenetic signal in continuous functional traits of the ferns of Japan.  
**`r s_table("phy-sig-binary")`**. Phylogenetic signal in quantitative (binary) functional traits of the ferns of Japan.  
**`r s_table("lrt-null")`**. Likelihood ratio test (LRT) between full model and null model (model only including the spatial Matérn correlation matrix).  
**`r s_table("aic")`**. Model fit as measured with conditional Akaike Information Criterion (cAIC).  
**`r s_figure("sampling-curve")`**. Species collection curve for the ferns of Japan.  
**`r s_figure("grain-size")`**. Effect of grain size on sampling redundancy.    
**`r s_figure("envir")`**. Reproductive mode and environmental data in 0.2 degree grid cells across Japan.  
**`r s_figure("tree")`**. Time tree of native Japanese ferns and lycophytes, excluding hybrids.  
**`r s_figure("abundance")`**. Observed number of specimens and sampling redundancy per 0.2° grid cell in the ferns of Japan.  
**`r s_figure("raw-lat")`**. Raw taxonomic diversity, phylogenetic diversity, and functional diversity of the ferns of Japan plotted by latitude.  
**`r s_figure("raw-biplot")`**. Relationships between observed functional and phylogenetic diversity and taxonomic richness in the ferns of Japan.  
**`r s_figure("endemism-restricted")`**. Phylogenetic endemism of the ferns of Japan measured using CANAPE (categorical analysis of neo- and paleo-endemism), restricted dataset including only taxa endemic to Japan.  
**`r s_figure("k-plot")`**. Selection of *k* for bioregions.  
**`r s_figure("dendrogram")`**. Bioregion dendrograms.  
**`r s_figure("genus-map")`**. Map of species richness for selected genera of ferns of Japan.  
**`r s_figure("apo-genus")`**. Rate of apomixis in genera of native, non-hybrid Japanese ferns with > 1 apomictic taxon.  
**`r s_figure("modelfit-not-ult")`**. Relationship between phylogenetic diversity (PD) and relative PD and % apomictic taxa, models inferred using phylogeny with branchlengths in units of expected genetic change (untransformed ML tree).  
**`r s_figure("bioregions-env")`**. Scatterplots of grid cell membership in taxonomic or phylogenetic bioregions arranged by mean annual temperature and annual precipitation.  
**`r s_figure("protected-map")`**. Maps showing overlap of grid-cells with significantly high biodiversity for the ferns of Japan and protected areas.  
**`r s_figure("deer-map")`**. Maps showing overlap of grid-cells with significantly high biodiversity for the ferns of Japan and distribution of Japanese deer (*Cervus nippon*).

# References
<!-- Insert <div id="refs"></div> to place the refs here instead of at the end -->
<div id="refs"></div>

@@

**`r figure("raw-diversity")`**. (A) Raw taxonomic diversity (B), phylogenetic diversity, and (C) functional diversity of the ferns of Japan. For (b) and (c), raw branchlengths were transformed to relative values by dividing each branchlength by the sum of all branchlengths.
